DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics

### 2016-12-06 2015CompleteDerivation/completeDerivation.tex
Comments from Christine on complex vector space

Suppose we have a state $\mathcal{c}_1$ for an amount $a_m$ of classical material. Suppose we assigned a numeric identifier to each particle of the material. This identifier would be just a dummy label we assign, not a physically quantity. As the classical material is homogeneous, permutations between classical particles within the same state have no effect. Therefore the ordering given by the identifier does not matter. Identifying a classical particle within a composite state is equivalent to picking a number from a random variable $A$ uniformly distributed over an interval of size $a_m$.

Suppose we have another state $\mathcal{c}_2 = \tau(a) \mathcal{c}_1$ where the only difference is an increase/decrease of the amount of material by a factor $a$. The random variable $B$ that identifies a particle of $\mathcal{c}_2$ is now a uniform distribution $B=aA$ where the range is the new amount of material: there are now proportionally more/less labels to assign. This is also the only possible transformation that can link the two uniform random variables $A$ and $B$. Therefore increase/decrease of material is the only operation that is going to affect the labeling of the fragments that is captured by the state. In other words: the classical state does not care about the permutation of the fragments, but it cares how many fragments are there.

@@

If the amount of material is connected to the area of two uniform distributions, it should combine like the variance of a random variable. And in fact it does. Suppose $X$ and $Y$ are two random variables,
\begin{equation}\label{eqn:correlation}
\sigma^2_{X+Y} = \sigma^2_{X} + \sigma^2_{Y} + 2 \, \sigma_{X} \sigma_{Y} \rho_{X,Y}
\end{equation}
where $\sigma^2$ is the variance, $\sigma$ is the standard deviation, and $\rho_{X,Y}=COV(X,Y)/\sigma_{X} \sigma_{Y}$ is the Pearson correlation coefficient. As $-1\leq\rho_{X,Y}\leq+1$, the total variance can be greater or smaller than the sum of the variance of the components. Now suppose $\psi$ is the state for a unit amount of irreducible material (i.e. a quantum particle). We can write $c \psi = (c_1 + c_2) \psi = c_1 \psi + c_2 \psi$. We have:
\begin{equation}\label{eqn:correlation_state}
|c_1+c_2|^2=|c_1|^2 + |c_2|^2 + 2 |c_1||c_2|\cos(\theta_{c_2} - \theta_{c_1})
\end{equation}
The two expressions are formally identical and it shouldn't come as a surprise. The more material we have, the more labels for the fragment configurations to choose from. $-1\leq\cos(\theta_{c_2} - \theta_{c_1})\leq+1$ takes the place of the Pearson correlation coefficient. Quantum interference can then be understood in terms of the combination of correlated/anti-correlated distributions.

@@

If quantum states are uniform distributions over fragment configurations, we should see a hint of that when computing distributions over observables. And in fact we do. Suppose we have $Q\psi = q \psi$, an eigenstate of some observable. Then it is also an eigenstate of the infinitesimal transformation of the conjugate variable $1+\frac{Qdp}{\imath\hbar}$. Since it has a symmetry over $p$, the distribution is uniform on that parameter.

If quantum particle states are really just unit amount of material uniformly distributed, we should be able to calculate the Shannon entropy associated with them. Using our dummy variables $I(\psi_i) = - \int \rho ln (\rho) dA dB = - \int 1 ln (1) dA dB = 0$. For a quantum mixed state $\tilde{\rho} = \sum \tilde{\rho}_i |\psi_i\rangle \langle \psi_i |$, we have $I(\tilde{\rho}) = - \sum \tilde{\rho}_i ln \tilde{\rho}_i - \tilde{\rho}_i I(\psi_i) = - \sum \tilde{\rho}_i ln \tilde{\rho}_i = -tr(\tilde{\rho} ln \tilde{\rho})$. And in fact this is the Von Neumann entropy used in quantum information, that implicitly assumes all states provide the same amount of entropy (i.e. zero).

@@

In other words, not only our simple model is consistent, but it has predictive power: it provides intuition that is well supported by these findings. It is not ``just a mathematical trick".

>>>>

Suppose we have a state $\mathcal{c}_1$ for an amount $a_m$ of classical material. Suppose we assigned a numeric identifier to each particle of the material. This identifier would be just a dummy label we assign, not a physical quantity. As the classical material is homogeneous, permutations between classical particles within the same state have no effect. Therefore the ordering given by the identifier does not matter. Identifying a classical particle within a composite state is equivalent to picking a number from a random variable $A$ uniformly distributed over an interval of size $a_m$.

Suppose we have another state $\mathcal{c}_2 = \tau(a) \mathcal{c}_1$ where the only difference is an increase/decrease of the amount of material by a factor $a$. The random variable $B$ that identifies a particle of $\mathcal{c}_2$ is now a uniform distribution $B=aA$ where the range is the new amount of material: there are now proportionally more/less labels to assign. This is also the only possible transformation that can link the two uniform random variables $A$ and $B$. Therefore increase/decrease of material is the only operation that is captured by the state among those that affect the labeling of the fragments. In other words: the classical state does not care about the permutation of the fragments, but it cares how many fragments are there.

@@

If the amount of material is connected to the area of two uniform distributions, it should combine like the variance of a random variable. And in fact it does. Suppose $X$ and $Y$ are two random variables,
\begin{equation}\label{eqn:correlation}
\sigma^2_{X+Y} = \sigma^2_{X} + \sigma^2_{Y} + 2 \, \sigma_{X} \sigma_{Y} \rho_{X,Y}
\end{equation}
where $\sigma^2$ is the variance, $\sigma$ is the standard deviation, and $\rho_{X,Y}=COV(X,Y)/\sigma_{X} \sigma_{Y}$ is the Pearson correlation coefficient. As $-1\leq\rho_{X,Y}\leq+1$, the total variance can be greater or smaller than the sum of the variance of the components. Now suppose $\psi$ is the state for a unit amount of irreducible material (i.e. a quantum particle). We can write $c \psi = (c_1 + c_2) \psi = c_1 \psi + c_2 \psi$. We have:
\begin{equation}\label{eqn:correlation_state}
|c_1+c_2|^2=|c_1|^2 + |c_2|^2 + 2 |c_1||c_2|\cos(\theta_{c_2} - \theta_{c_1})
\end{equation}
The two expressions are formally identical and it shouldn't come as a surprise. The more material we have, the more labels for the fragment configurations to choose from, the bigger the spread of the random variables that assign the labels. $-1\leq\cos(\theta_{c_2} - \theta_{c_1})\leq+1$ takes the place of the Pearson correlation coefficient. Quantum interference can then be understood in terms of the combination of correlated/anti-correlated distributions.

@@

If quantum states are uniform distributions over fragment configurations, we should see a hint of that when computing distributions over observables. And in fact we do. Suppose we have $Q\psi = q \psi$, an eigenstate of some observable. Then it is also an eigenstate of the infinitesimal transformation of the conjugate variable $1+\frac{Qdp}{\imath\hbar}$. Since it has a symmetry over $p$, the distribution has to be equal at each value of $p$: the distribution is uniform.

If quantum particle states are really just unit amounts of material uniformly distributed, we should be able to calculate the Shannon entropy associated with them. Using our dummy variables, $I(\psi) = - \int \rho \ln (\rho) dA dB = - \int 1 ln (1) dA dB = 0$. For a quantum mixed state $\tilde{\rho} = \sum \tilde{\rho}_i |\psi_i\rangle \langle \psi_i |$, we have $I(\tilde{\rho}) = - \sum \tilde{\rho}_i \ln \tilde{\rho}_i - \tilde{\rho}_i I(\psi_i) = - \sum \tilde{\rho}_i \ln \tilde{\rho}_i = -tr(\tilde{\rho} \ln \tilde{\rho})$. And in fact this is the Von Neumann entropy used in quantum information, that implicitly assumes all states provide the same amount of entropy (i.e. zero).

@@

In other words, our simple model is not only consistent with quantum mechanics but it has predictive power: it provides intuition that is well supported by these findings. It is not ``just a mathematical trick".
